---
title: "Model Scores"
output: html_notebook
---

👻👻👻 Boo! I am a ghost!! 👻👻👻 

But I am harmless. 👻

```{r}
library(rsample)      # data splitting 
library(randomForest) # basic implementation
library(ranger)       # a faster implementation of randomForest
library(caret)        # an aggregator package for performing many machine learnin
library(h2o)          # an extremely fast java-based platform
library(dplyr)
library(magrittr)

r2_general <-function(preds,actual){ 
  return(1- sum((preds - actual) ^ 2)/sum((actual - mean(actual))^2))
}

smear <- read.csv(file = '/home/laatopi/Documents/SmearTest/test_data.csv')
smear <- within(smear, rm(X, SoilWatCont))

smear_split <- initial_split(smear, prop = .7)
smear_train <- training(smear_split)
smear_test  <- testing(smear_split)

valid_split <- initial_split(smear_train, .8)
smear_train_2 <- analysis(valid_split)

smear_valid <- assessment(valid_split)
x_test <- smear_valid[setdiff(names(smear_valid), "NEP")]
y_test <- smear_valid$NEP

```


```{r}
OOB_RMSE <- vector(mode = "numeric", length = 1)

for(i in seq_along(OOB_RMSE)) {

  optimal_ranger <- ranger(
    formula         = NEP ~ ., 
    data            = smear_train, 
    num.trees       = 500,
    mtry            = 3,
    min.node.size   = 5,
    sample.fraction = .8,
    importance      = 'impurity'
  )
  
  OOB_RMSE[i] <- (optimal_ranger$prediction.error)
}

hist(OOB_RMSE, breaks = 20)
```
Variable importance is measured by recording the decrease in MSE each time a variable is used as a node split in a tree. The remaining error left in predictive accuracy after a node split is known as node impurity and a variable that reduces this impurity is considered more important than those variables that do not. Consequently, we accumulate the reduction in MSE for each variable across all the trees and the variable with the greatest accumulated impact is considered the more important, or impactful.
```{r}
sqrt(optimal_ranger$variable.importance) %>% 
  broom::tidy() %>%
  dplyr::arrange(desc(x)) %>%
  dplyr::top_n(10) %>%
  ggplot(aes(reorder(names, x), x)) +
  geom_col() +
  coord_flip() +
  ggtitle("Most important variables)")
```

```{r}
h2o.no_progress()
h2o.init(max_mem_size = "5g")
h2o.removeAll()
```

```{r}
y <- "NEP"
x <- setdiff(names(smear_train), y)

# turn training set into h2o object
train.h2o <- as.h2o(smear_train)

hyper_grid.h2o <- list(
  ntrees      = seq(200, 500, by = 150),
  mtries      = seq(1, 10, by = 1),
  max_depth   = seq(20, 40, by = 5),
  min_rows    = seq(1, 5, by = 2),
  nbins       = seq(10, 30, by = 5),
  sample_rate = c(.55, .632, .75)
)

# random grid search criteria
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.005,
  stopping_rounds = 10,
  max_runtime_secs = 25*60
  )

# build grid search 
random_grid <- h2o.grid(
  algorithm = "randomForest",
  grid_id = "rf_grid2",
  x = x, 
  y = y, 
  training_frame = train.h2o,
  hyper_params = hyper_grid.h2o,
  search_criteria = search_criteria
  )

# collect the results and sort by our model performance metric of choice
grid_perf2 <- h2o.getGrid(
  grid_id = "rf_grid2", 
  sort_by = "mse", 
  decreasing = FALSE
  )
print(grid_perf2)

best_model_id <- grid_perf2@model_ids[[1]]
best_model <- h2o.getModel(grid_perf2@model_ids[[1]])
```


```{r warning=FALSE}
# Now let’s evaluate the model performance on a test set
n = 30
MSE  <- vector(mode = "numeric", length = n)
RMSE <- vector(mode = "numeric", length = n)
MAE  <- vector(mode = "numeric", length = n)
R2   <- vector(mode = "numeric", length = n)
R2_adjusted <- vector(mode = "numeric", length = n)

cMSE  <- vector(mode = "numeric", length = n)
cRMSE <- vector(mode = "numeric", length = n)

basic_rf <- randomForest(
  formula = NEP ~ .,
  data    = smear_train
)

for(i in seq_along(MSE)) {
  
  smear_split <- initial_split(smear, prop = .5)
  smear_test  <- testing(smear_split)
  
  smear_test.h2o <- as.h2o(smear_test)
  best_model_perf <- h2o.performance(model = best_model, newdata = smear_test.h2o)
  RMSE[i] <- h2o.rmse(best_model_perf)
  MSE[i]  <- h2o.mse(best_model_perf)
  pred_h2o <- as.data.frame(predict(best_model, smear_test.h2o))$predict
  
  cMSE[i] <- mean((smear_test$NEP - pred_h2o)^2)
  cRMSE[i] <- sqrt(mean((smear_test$NEP - pred_h2o)^2))

  MAE[i] <- sum(abs(smear_test$NEP - pred_h2o))/ length(smear_test$NEP)
  R2[i] <- cor(smear_test$NEP, pred_h2o)^2
  k = length(pred_h2o)
  p =  10
  R2_adjusted[i] <- 1 - (((1 - R2[i]) * (k - 1)) / (k - p - 1))
  
}
```

```{r}
library(tidyr)
library(ggplot2)
l <- data.frame(MSE, RMSE, MAE, R2, R2_adjusted)

ggplot(gather(l), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free_x')
```

```{r}
high_points <- read.csv(file = "/home/laatopi/Documents/SmearTest/high_points.csv", header = TRUE)[, -1]
low_points  <- read.csv(file = "/home/laatopi/Documents/SmearTest/low_points.csv", header = TRUE)[, -1]
```

```{r}
library(lubridate)
names(high_points)[names(high_points) == 'HYY_EDDY233.u_star'] <- 'FricVel'
names(high_points)[names(high_points) == 'HYY_META.T336'] <- 'AirTemp'
names(high_points)[names(high_points) == 'HYY_META.tsoil_B2'] <- 'SoilTempB'
names(high_points)[names(high_points) == 'HYY_META.tsoil_A'] <- 'SoilTempA'
names(high_points)[names(high_points) == 'HYY_META.RHTd'] <- 'RelHum'
names(high_points)[names(high_points) == 'HYY_META.PAR2'] <- 'PPFD'
names(high_points)[names(high_points) == 'HYY_META.wsoil_B2'] <- 'SoilWatCont'
names(high_points)[names(high_points) == 'HYY_META.diffPAR'] <- 'PPFDdiff'
names(high_points)[names(high_points) == 'HYY_EDDY233.NEE'] <- 'NEP'

names(low_points)[names(low_points) == 'HYY_EDDY233.u_star'] <- 'FricVel'
names(low_points)[names(low_points) == 'HYY_META.T336'] <- 'AirTemp'
names(low_points)[names(low_points) == 'HYY_META.tsoil_B2'] <- 'SoilTempB'
names(low_points)[names(low_points) == 'HYY_META.tsoil_A'] <- 'SoilTempA'
names(low_points)[names(low_points) == 'HYY_META.RHTd'] <- 'RelHum'
names(low_points)[names(low_points) == 'HYY_META.PAR2'] <- 'PPFD'
names(low_points)[names(low_points) == 'HYY_META.wsoil_B2'] <- 'SoilWatCont'
names(low_points)[names(low_points) == 'HYY_META.diffPAR'] <- 'PPFDdiff'
names(low_points)[names(low_points) == 'HYY_EDDY233.NEE'] <- 'NEP'

high.h2o <- as.h2o(within(high_points, rm(Time)))
pred_high <- as.data.frame(predict(best_model, high.h2o))$predict

low.h2o <- as.h2o(within(low_points, rm(Time)))
pred_low <- as.data.frame(predict(best_model, low.h2o))$predict

df1 <- data.frame(number = pred_high, time = hour(high_points$Time))
df2 <- data.frame(number = pred_low, time = hour(low_points$Time))

df1$group <- 'High Prediction'
df2$group <- 'Low Prediction'

df <- rbind(df1, df2)

df3 <- data.frame(number = high_points$NEP, time = hour(high_points$Time))
df4 <- data.frame(number = low_points$NEP, time = hour(low_points$Time))
df3$group <- 'High Actual'
df4$group <- 'Low Actual'

df5 <- rbind(df1, df2, df3, df4)

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df2 <- data_summary(df5, varname="number", 
                    groupnames=c("time", "group"))

p <- ggplot(df2, aes(x=as.factor(time), y=number, group=group, color=group)) + 
  geom_line(position = position_dodge(0.5)) +
  geom_point(position = position_dodge(0.5))+
  geom_errorbar(aes(ymin=number-sd, ymax=number+sd), width=.2,
                 position=position_dodge(0.5))

p + labs(title="Actual and Predicted NEE values for High and Low Aerosol", x = "Hour" , y = "NEE", color="Type")+
   theme_classic() +
   scale_color_manual(values=c('#FF3333','#FF9933', '#3352FF', '#33D1FF')) +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
   theme(panel.grid.major.y = element_line(color = "gray",
                                        size = 0.5,
                                        linetype = 2))
```

```{r warning=FALSE}
library(iml)

pred <- function(model, newdata)  {
  results <- as.data.frame(h2o.predict(model, as.h2o(newdata)))
  return(results[[3L]])
}

predictor_h <- Predictor$new(
  model = best_model, 
  data = high_points[which(names(high_points) != c("Time"))], 
  y = "NEP", 
  predict.fun = pred
)

eff <- FeatureEffects$new(predictor_h, method = "ale")
eff$plot()

predictor_l <- Predictor$new(
  model = best_model, 
  data = low_points[which(names(low_points) != c("Time"))], 
  y = "NEP", 
  predict.fun = pred
)

eff <- FeatureEffects$new(predictor_l, method = "ale")
eff$plot()

imp.h <- FeatureImp$new(predictor_h, loss = "mse", n.repetitions = 12)
imp.l <- FeatureImp$new(predictor_l, loss = "mse", n.repetitions = 12)
# plot output

p1 <- plot(imp.h) + ggtitle("High")
p2 <- plot(imp.l) + ggtitle("Low")
gridExtra::grid.arrange(p1, p2, nrow = 1)
```
The part with the partial models :D
```{r}
OOB_RMSE <- vector(mode = "numeric", length = 10)
R2 <- vector(mode = "numeric", length = 10)
R2_2 <- vector(mode = "numeric", length = 10)

kerroshamppari <- data.frame(matrix(ncol=0, nrow=10))
low_R2 <- data.frame(matrix(ncol=0, nrow=10))
high_R2 <- data.frame(matrix(ncol=0, nrow=10))

for(i in c(3, 2, 4:10)) {

  var <- colnames(smear_train[, c(1, i)])[2]
  
  for(j in seq_along(OOB_RMSE)) {
    smear_split <- initial_split(smear, prop = .7)
    smear_train <- training(smear_split)
    smear_test  <- testing(smear_split)
    
    optimal_ranger <- ranger(
      formula         = NEP ~ .,
      data            = smear_train[, c(1, i)],
      num.trees       = 500,
      mtry            = 1,
      min.node.size   = 5,
      sample.fraction = .8,
      importance      = 'impurity'
    )
    
    predictions <- predict(optimal_ranger, smear_test[var])
    
    reg_score <- cor(smear_test$NEP, predictions$predictions)^2
    
    predictions <- predict(optimal_ranger, low_points[var])
  
    l_score <- cor(low_points$NEP, predictions$predictions)^2
    
    predictions <- predict(optimal_ranger, high_points[var])
    
    h_score <- cor(high_points$NEP, predictions$predictions)^2
    
    OOB_RMSE[j] <- reg_score
    R2[j] <- l_score
    R2_2[j] <- h_score
  }

  kerroshamppari[var] <- OOB_RMSE
  low_R2[var] <- R2
  high_R2[var] <- R2_2
}
```

Plotting

-> -> ->
  -> -> ->
    -> -> ->
  -> -> ->
-> -> ->
```{r}
plot_for_all <- function() {
  results <- sapply(kerroshamppari, function(x) c("SD" = sd(x), "Mean" = mean(x)))
  results <- data.frame(t(results))
  results <- tibble::rownames_to_column(results, "Var")
  
  p <- ggplot(results, aes(x=reorder(Var, Mean), y=Mean, fill=Var)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2,
                   position=position_dodge(.9)) 
  p + labs(title="Mean R2 of Random Forest trained with single Var, n=2528", x="Variable", y = "Mean RMSE")+
     theme_classic() +
     theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
     theme(legend.position = "none")
}

plot_for_all()
```

```{r}
plot_for_lowNhigh <- function() {

  results_low <- sapply(low_R2, function(x) c("SD" = sd(x), "Mean" = mean(x))) %>% 
    t() %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column("Var")
  
  results_high <- sapply(high_R2, function(x) c("SD" = sd(x), "Mean" = mean(x))) %>% 
    t() %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column("Var")
  
  p1 <- ggplot(results_low, aes(x=reorder(Var, Mean), y=Mean, fill=Var)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2,
                   position=position_dodge(.9)) 
  p1 <- p1 + labs(title="Mean R2 of Low, n=341", x="Variable", y = "Mean RMSE")+
     theme_classic() +
     theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
     theme(legend.position = "none")
  
  p2 <- ggplot(results_high, aes(x=reorder(Var, Mean), y=Mean, fill=Var)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2,
                   position=position_dodge(.9)) 
  p2 <- p2 + labs(title="Mean R2 of High, n=344", x="Variable", y = "Mean RMSE")+
     theme_classic() +
     theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
     theme(legend.position = "none")
  
  
  gridExtra::grid.arrange(p1, p2, nrow = 1)
}

plot_for_lowNhigh()
```


```{r}
OOB_RMSE <- vector(mode = "numeric", length = 10)
R2 <- vector(mode = "numeric", length = 10)
R2_2 <- vector(mode = "numeric", length = 10)

kerroshamppari <- data.frame(matrix(ncol=0, nrow=10))
low_R2 <- data.frame(matrix(ncol=0, nrow=10))
high_R2 <- data.frame(matrix(ncol=0, nrow=10))
for(i in c(2, 4:10)) {

  var <- colnames(smear_train[, c(1, 3, i)])[3]
  
  for(j in seq_along(OOB_RMSE)) {
    smear_split <- initial_split(smear, prop = .7)
    smear_train <- training(smear_split)
    smear_test  <- testing(smear_split)
    
    optimal_ranger <- ranger(
      formula         = NEP ~ .,
      data            = smear_train[, c(1, 3, i)],
      num.trees       = 500,
      mtry            = 2,
      min.node.size   = 5,
      sample.fraction = .8,
      importance      = 'impurity'
    )
    
    predictions <- predict(optimal_ranger, smear_test[c("PPFD",var)])
    
    reg_score <- cor(smear_test$NEP, predictions$predictions)^2
    
    predictions <- predict(optimal_ranger, low_points[c("PPFD",var)])
  
    l_score <- cor(low_points$NEP, predictions$predictions)^2
    
    predictions <- predict(optimal_ranger, high_points[c("PPFD",var)])
    
    h_score <- cor(high_points$NEP, predictions$predictions)^2
    
    OOB_RMSE[j] <- reg_score
    R2[j] <- l_score
    R2_2[j] <- h_score
  }

  kerroshamppari[var] <- OOB_RMSE
  low_R2[var] <- R2
  high_R2[var] <- R2_2
}

plot_for_all()
plot_for_lowNhigh()
```

```{r}
print("DONE!")
```