---
title: "Model Scores"
output: html_notebook
---

👻👻👻 Boo! I am a ghost!! 👻👻👻 

But I am harmless. 👻

```{r}
library(rsample)      # data splitting 
library(randomForest) # basic implementation
library(ranger)       # a faster implementation of randomForest
library(caret)        # an aggregator package for performing many machine learnin
library(h2o)          # an extremely fast java-based platform
library(dplyr)
library(magrittr)
library(iml)
library(tidyr)
library(ggplot2)
library(lubridate)
library(grid)
library(gridExtra)
library(cowplot)


smear_2012_2019 <- within(read.csv(file = '/home/laatopi/Documents/SmearTest/test_data.csv'), rm(X, SoilWatCont))
smear_2008_2011 <- within(read.csv(file = '/home/laatopi/Documents/SmearTest/smear2008-11.csv'), rm(X, Time))

high_points <- read.csv(file = "/home/laatopi/Documents/SmearTest/high_points.csv", header = TRUE)[, -1]
low_points  <- read.csv(file = "/home/laatopi/Documents/SmearTest/low_points.csv", header = TRUE)[, -1]

names(high_points)[names(high_points) == 'HYY_EDDY233.u_star'] <- 'FricVel'
names(high_points)[names(high_points) == 'HYY_META.T336'] <- 'AirTemp'
names(high_points)[names(high_points) == 'HYY_META.tsoil_B2'] <- 'SoilTempB'
names(high_points)[names(high_points) == 'HYY_META.tsoil_A'] <- 'SoilTempA'
names(high_points)[names(high_points) == 'HYY_META.RHTd'] <- 'RelHum'
names(high_points)[names(high_points) == 'HYY_META.PAR2'] <- 'PPFD'
names(high_points)[names(high_points) == 'HYY_META.wsoil_B2'] <- 'SoilWatCont'
names(high_points)[names(high_points) == 'HYY_META.diffPAR'] <- 'PPFDdiff'
names(high_points)[names(high_points) == 'HYY_EDDY233.NEE'] <- 'NEP'

names(low_points)[names(low_points) == 'HYY_EDDY233.u_star'] <- 'FricVel'
names(low_points)[names(low_points) == 'HYY_META.T336'] <- 'AirTemp'
names(low_points)[names(low_points) == 'HYY_META.tsoil_B2'] <- 'SoilTempB'
names(low_points)[names(low_points) == 'HYY_META.tsoil_A'] <- 'SoilTempA'
names(low_points)[names(low_points) == 'HYY_META.RHTd'] <- 'RelHum'
names(low_points)[names(low_points) == 'HYY_META.PAR2'] <- 'PPFD'
names(low_points)[names(low_points) == 'HYY_META.wsoil_B2'] <- 'SoilWatCont'
names(low_points)[names(low_points) == 'HYY_META.diffPAR'] <- 'PPFDdiff'
names(low_points)[names(low_points) == 'HYY_EDDY233.NEE'] <- 'NEP'
```


Variable importance is measured by recording the decrease in MSE each time a variable is used as a node split in a tree. The remaining error left in predictive accuracy after a node split is known as node impurity and a variable that reduces this impurity is considered more important than those variables that do not. Consequently, we accumulate the reduction in MSE for each variable across all the trees and the variable with the greatest accumulated impact is considered the more important, or impactful.

```{r}
h2o.no_progress()
h2o.init(max_mem_size = "8g")
h2o.removeAll()
```

Random Forest
```{r}
tune_random_forest <- function(data) {
  
  y <- "NEP"
  x <- setdiff(names(data), y)
  
  # turn training set into h2o object
  train.h2o <- as.h2o(data)
  
  splits <- h2o.splitFrame(
    data = as.h2o(data), 
    ratios = c(0.6,0.2),   ## only need to specify 2 fractions, the 3rd is implied
    destination_frames = c("train.hex", "valid.hex", "test.hex")
  )
  
  train <- splits[[1]]
  valid <- splits[[2]]
  test  <- splits[[3]]
  
  hyper_params <- list(
    ntrees      = seq(200, 500, by = 150),
    mtries      = seq(1, 10, by = 1),
    max_depth   = seq(20, 40, by = 5),
    min_rows    = seq(1, 5, by = 2),
    nbins       = seq(10, 30, by = 5),
    sample_rate = c(.55, .632, .75)
  )
  
  # random grid search criteria
  search_criteria <- list(
    strategy = "RandomDiscrete",
    stopping_metric = "mse",
    stopping_tolerance = 0.005,
    stopping_rounds = 10,
    max_runtime_secs = 45*60,
    max_models = 5                 
    )
  
  # build grid search
  grid <- h2o.grid(
    algorithm = "randomForest",
    grid_id = "rf_grid2",
    x = x, 
    y = y, 
    training_frame = train, 
    validation_frame = valid,
    hyper_params = hyper_params,
    search_criteria = search_criteria
    )
  
  # collect the results and sort by our model performance metric of choice
  grid_perf <- h2o.getGrid(
    grid_id = "rf_grid2",
    sort_by = "mse",
    decreasing = FALSE
  )

  best_model_id <- grid_perf@model_ids[[1]]
  best_model <- h2o.getModel(grid_perf@model_ids[[1]])
  h2o.removeAll()
  return(best_model)
}
```

GBM
```{r}
tune_gbm <- function(data) {
  
  y <- "NEP"
  x <- setdiff(names(data), y)
  
  splits <- h2o.splitFrame(
    data = as.h2o(data), 
    ratios = c(0.6,0.2),   ## only need to specify 2 fractions, the 3rd is implied
    destination_frames = c("train.hex", "valid.hex", "test.hex")
  )
  
  train <- splits[[1]]
  valid <- splits[[2]]
  test  <- splits[[3]]
  
  hyper_params = list( 
    max_depth                = seq(1,29,2),                                      
    sample_rate              = seq(0.2, 1, 0.01),                                             
    col_sample_rate          = seq(0.2, 1, 0.01),                                         
    col_sample_rate_per_tree = seq(0.2, 1, 0.01),                                
    col_sample_rate_change_per_level = seq(0.9, 1.1, 0.01),                      
    min_rows = 2^seq(0,log2(nrow(train))-1,1),                                 
    nbins = 2^seq(4, 10, 1),                                                     
    nbins_cats = 2^seq(4, 12, 1),                                                
    min_split_improvement = c(0,1e-8, 1e-6, 1e-4),
    ntrees = seq(200, 1000, by = 150),                                                           
    histogram_type = c("UniformAdaptive","QuantilesGlobal","RoundRobin")       
  )
  
  search_criteria = list(
    strategy = "RandomDiscrete",      
    stopping_metric = "mse",
    stopping_tolerance = 0.005,
    stopping_rounds = 10,
    max_runtime_secs = 20*60,         
    max_models = 1                 
  )
  
  grid <- h2o.grid(
    algorithm = "gbm",
    grid_id = "final_grid", 
    x = x, 
    y = y, 
    training_frame = train, 
    validation_frame = valid,
    hyper_params = hyper_params,
    search_criteria = search_criteria,
    score_tree_interval = 10,
    learn_rate = 0.05,                                                         
    learn_rate_annealing = 0.99
  )
  
  grid_perf <- h2o.getGrid(
    grid_id = "final_grid", 
    sort_by = "mse", 
    decreasing = FALSE
  )
  
  best_model_id <- grid_perf@model_ids[[1]]
  best_model <- h2o.getModel(grid_perf@model_ids[[1]])
  h2o.removeAll()
  return(best_model)
}
```

```{r warning=FALSE}
test_model_perf <- function(model, data, n) {
  n = n
  MSE  <- vector(mode = "numeric", length = n)
  RMSE <- vector(mode = "numeric", length = n)
  MAE  <- vector(mode = "numeric", length = n)
  R2   <- vector(mode = "numeric", length = n)
  R2_adjusted <- vector(mode = "numeric", length = n)
  
  cMSE  <- vector(mode = "numeric", length = n)
  cRMSE <- vector(mode = "numeric", length = n)
  
  for(i in seq_along(MSE)) {
    
    split <- initial_split(data, prop = .5)
    test  <- testing(smear_split)
    
    test.h2o <- as.h2o(test)
    model_perf <- h2o.performance(model = model, newdata = test.h2o)
    RMSE[i] <- h2o.rmse(model_perf)
    MSE[i]  <- h2o.mse(model_perf)
    
    pred_h2o <- as.data.frame(predict(model, test.h2o))$predict

    MAE[i] <- sum(abs(test$NEP - pred_h2o))/ length(test$NEP)
    R2[i] <- cor(test$NEP, pred_h2o)^2
    k = length(pred_h2o)
    p = 9
    R2_adjusted[i] <- 1 - (((1 - R2[i]) * (k - 1)) / (k - p - 1))
  }
  
  l <- data.frame(MSE, RMSE, MAE, R2, R2_adjusted)

  ggplot(gather(l), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free_x')
}
```

```{r}
create_hilo_pred_frame <- function(model, high, low) {
  
  high.h2o <- as.h2o(within(high, rm(Time)))
  pred_high <- as.data.frame(predict(model, high.h2o))$predict
  
  low.h2o <- as.h2o(within(low, rm(Time)))
  pred_low <- as.data.frame(predict(model, low.h2o))$predict
  
  df1 <- data.frame(number = pred_high, time = hour(high$Time))
  df2 <- data.frame(number = pred_low, time = hour(low$Time))
  
  df1$group <- 'High Prediction'
  df2$group <- 'Low Prediction'
  
  df <- rbind(df1, df2)
  
  df3 <- data.frame(number = high$NEP, time = hour(high$Time))
  df4 <- data.frame(number = low$NEP, time = hour(low$Time))
  df3$group <- 'High Actual'
  df4$group <- 'Low Actual'
  
  df5 <- rbind(df1, df2, df3, df4)
  
  varname <- "number"
  groupnames <- c("time", "group")
  
  require(plyr)
  
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  
  data_sum <- ddply(df5, groupnames, .fun=summary_func, varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}


plot_hi_lo <- function(df2, title_text) {

  p <- ggplot(df2, aes(x=as.factor(time), y=number, group=group, color=group)) + 
    geom_line(position = position_dodge(0.5)) +
    geom_point(position = position_dodge(0.5))+
    geom_errorbar(aes(ymin=number-sd, ymax=number+sd), width=.2,
                   position=position_dodge(0.5))
  
  p <- p + labs(title = title_text, x = "Hour" , y = "NEE", color="Type")+
     theme_classic() +
     scale_color_manual(values=c('#FF3333','#FF9933', '#3352FF', '#33D1FF')) +
     scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
     theme(panel.grid.major.y = element_line(color = "gray",
                                          size = 0.5,
                                          linetype = 2))
  return(p)
}
```

```{r warning=FALSE}
pred <- function(model, newdata)  {
  results <- as.data.frame(h2o.predict(model, as.h2o(newdata)))
  return(results[[3L]])}

predict_feature_effects <- function(model) {

  predictor_h <- Predictor$new(
    model = model, 
    data = high_points[which(names(high_points) != c("Time"))], 
    y = "NEP", 
    predict.fun = pred
  )
  
  eff <- FeatureEffects$new(predictor_h, method = "ale")
  eff$plot()
  
  predictor_l <- Predictor$new(
    model = model, 
    data = low_points[which(names(low_points) != c("Time"))], 
    y = "NEP", 
    predict.fun = pred
  )
  
  eff <- FeatureEffects$new(predictor_l, method = "ale")
  eff$plot()
  
  imp.h <- FeatureImp$new(predictor_h, loss = "mse", n.repetitions = 12)
  imp.l <- FeatureImp$new(predictor_l, loss = "mse", n.repetitions = 12)
  # plot output
  
  p1 <- plot(imp.h) + ggtitle("High")
  p2 <- plot(imp.l) + ggtitle("Low")
  
  return(gridExtra::grid.arrange(p1, p2, nrow = 1))
}
```
The part with the partial models :D


```{r}
plot_for_all <- function(all_R2) {
  results <- sapply(all_R2, function(x) c("SD" = sd(x), "Mean" = mean(x)))
  results <- data.frame(t(results))
  results <- tibble::rownames_to_column(results, "Var")
  
  p <- ggplot(results, aes(x=reorder(Var, Mean), y=Mean, fill=Var)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2,
                   position=position_dodge(.9)) 
  p + labs(title="Mean R2 of Random Forest trained with single Var, n=2528", x="Variable", y = "Mean RMSE")+
     theme_classic() +
     theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
     theme(legend.position = "none")
}
```

```{r}
plot_for_lowNhigh <- function(low_R2, high_R2, title) {

  results_low <- sapply(low_R2, function(x) c("SD" = sd(x), "Mean" = mean(x))) %>% 
    t() %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column("Var")
  
  results_high <- sapply(high_R2, function(x) c("SD" = sd(x), "Mean" = mean(x))) %>% 
    t() %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column("Var")
  
  p1 <- ggplot(results_low, aes(x=reorder(Var, Mean), y=Mean, fill=Var)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2,
                   position=position_dodge(.9)) 
  p1 <- p1 + labs(title="Mean R2 of Low", x="Variable", y = "Mean RMSE")+
     theme_classic() +
     theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
     theme(legend.position = "none") + 
     theme(panel.grid.major = element_line(colour = "gray", size = (0.2)), panel.grid.minor = element_line(size = (0.2), colour ="grey")) +
     ylim(0, 1)
  
  p2 <- ggplot(results_high, aes(x=reorder(Var, Mean), y=Mean, fill=Var)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2,
                   position=position_dodge(.9)) 
  p2 <- p2 + labs(title="Mean R2 of High", x="Variable", y = "Mean RMSE")+
     theme_classic() +
     theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
     theme(legend.position = "none") +
     theme(panel.grid.major = element_line(colour = "gray", size = (0.2)), panel.grid.minor = element_line(size = (0.2), colour ="grey")) +
     ylim(0, 1)
  
  return(grid.arrange(p1, p2, nrow = 1, top = textGrob(title, gp=gpar(fontsize=12,font=3))))
}
```

```{r}
# OOB_RMSE <- vector(mode = "numeric", length = 10)
# R2 <- vector(mode = "numeric", length = 10)
# R2_2 <- vector(mode = "numeric", length = 10)
# 
# all_R2 <- data.frame(matrix(ncol=0, nrow=10))
# low_R2 <- data.frame(matrix(ncol=0, nrow=10))
# high_R2 <- data.frame(matrix(ncol=0, nrow=10))
# 
# for(i in c(1:2, 4:10)) {
# 
#   var <- colnames(smear[, c(3, i)])[2]
#   k <- 1
# 
#   for(j in seq_along(OOB_RMSE)) {
#     
#     splits <- h2o.splitFrame(
#       data = as.h2o(smear[c("NEP", var)]), 
#       ratios = c(0.6,0.2),
#       destination_frames = c("train.hex", "valid.hex", "test.hex")
#     )
#     
#     train <- splits[[1]]
#     valid <- splits[[2]]
#     test  <- splits[[3]]    
#       
# 
#     curr_model <- do.call(h2o.gbm,
#         {
#           p <- best_model@parameters
#           p$model_id = NULL          ## do not overwrite the original grid model
#           p$training_frame = train      ## use the full dataset
#           p$validation_frame = NULL  ## no validation frame
#           p$nfolds = 5               ## cross-validation
#           p$x = var
#           p$y = y
#           p
#         }
#     )
#     
#     curr_model <- gbm
#     k <- k + 1
#     
#     predicted_y <- as.data.frame(predict(curr_model, as.h2o(smear_test[var])))$predict
#     reg_score <- cor(smear_test$NEP, predicted_y)^2
#     
#     predicted_y <- as.data.frame(predict(curr_model, as.h2o(low_points[var])))$predict
#     l_score <- cor(low_points$NEP, predicted_y)^2
#     
#     predicted_y <- as.data.frame(predict(curr_model, as.h2o(high_points[var])))$predict
#     h_score <- cor(high_points$NEP, predicted_y)^2
#     
#     OOB_RMSE[j] <- reg_score
#     R2[j] <- l_score
#     R2_2[j] <- h_score
#   }
# 
#   all_R2[var] <- OOB_RMSE
#   low_R2[var] <- R2
#   high_R2[var] <- R2_2
# }
```


```{r}
train_with_two <- function(data, model1, model2, n) {

  gbm_R2_all  <- vector(mode = "numeric", length = n)
  gbm_R2_l    <- vector(mode = "numeric", length = n)
  gbm_R2_h    <- vector(mode = "numeric", length = n)
  
  gbm_all_R2  <- data.frame(matrix(ncol=0, nrow = n))
  gbm_low_R2  <- data.frame(matrix(ncol=0, nrow = n))
  gbm_high_R2 <- data.frame(matrix(ncol=0, nrow = n))
  
  rf_R2_all   <- vector(mode = "numeric", length = n)
  rf_R2_l     <- vector(mode = "numeric", length = n)
  rf_R2_h     <- vector(mode = "numeric", length = n)
  
  rf_all_R2   <- data.frame(matrix(ncol=0, nrow = n))
  rf_low_R2   <- data.frame(matrix(ncol=0, nrow = n))
  rf_high_R2  <- data.frame(matrix(ncol=0, nrow = n))
    
  varnames <- c("AirTemp", "FricVel", "PPFDdiff", "RelHum", "SoilTempA", "SoilTempB", "DiffuseFract", "VaporPressureDeficit")

  for(var in varnames) {
    print(var)
    for(j in seq_along(gbm_R2_all)) {
  
      splits <- h2o.splitFrame(
        data = as.h2o(data[c("NEP", "PPFD", var)]), 
        ratios = c(0.6,0.2),
        destination_frames = c("train.hex", "valid.hex", "test.hex")
      )
      
      train <- splits[[1]]
      valid <- splits[[2]]
      test  <- splits[[3]]
      
      gbm <- do.call(h2o.gbm, {
        p <- model1@parameters
        p$model_id = NULL          ## do not overwrite the original grid model
        p$training_frame = train      ## use the full dataset
        p$validation_frame = NULL  ## no validation frame
        p$nfolds = 5               ## cross-validation
        p$x = c(var, "PPFD")
        p$y = "NEP"
        p
      })
      curr_model <- gbm
      
      predicted_y <- as.data.frame(predict(curr_model, test))$predict
      reg_score <- cor(as.data.frame(test)$NEP, predicted_y)^2
      
      predicted_y <- as.data.frame(predict(curr_model, as.h2o(low_points[c("PPFD", var)])))$predict
      l_score <- cor(low_points$NEP, predicted_y)^2
      
      predicted_y <- as.data.frame(predict(curr_model, as.h2o(high_points[c("PPFD", var)])))$predict
      h_score <- cor(high_points$NEP, predicted_y)^2
      
      gbm_R2_all[j] <- reg_score
      gbm_R2_l[j]   <- l_score
      gbm_R2_h[j]   <- h_score
      
      rf <- do.call(h2o.randomForest, {
        p <- model2@parameters
        p$model_id = NULL
        p$training_frame = train      ## use the full dataset
        p$validation_frame = NULL  ## no validation frame
        p$nfolds = 5               ## cross-validation
        p$x = c(var, "PPFD")
        p$y = "NEP"
        p$mtries = -1
        p
      })
      
      curr_model <- rf
      
      predicted_y <- as.data.frame(predict(curr_model, test))$predict
      reg_score <- cor(as.data.frame(test)$NEP, predicted_y)^2
      
      predicted_y <- as.data.frame(predict(curr_model, as.h2o(low_points[c("PPFD", var)])))$predict
      l_score <- cor(low_points$NEP, predicted_y)^2
      
      predicted_y <- as.data.frame(predict(curr_model, as.h2o(high_points[c("PPFD", var)])))$predict
      h_score <- cor(high_points$NEP, predicted_y)^2
      
      rf_R2_all[j] <- reg_score
      rf_R2_l[j]   <- l_score
      rf_R2_h[j]   <- h_score
      
    }
    h2o.removeAll()
    gbm_all_R2[var]  <- gbm_R2_all
    gbm_low_R2[var]  <- gbm_R2_l
    gbm_high_R2[var] <- gbm_R2_h
    
    rf_all_R2[var]  <- rf_R2_all
    rf_low_R2[var]  <- rf_R2_l
    rf_high_R2[var] <- rf_R2_h  

  }
  scores <- list("gbm_all" = gbm_all_R2, "gbm_low" = gbm_low_R2, "gbm_high" = gbm_high_R2, "rf_all" = rf_all_R2, "rf_low" = rf_low_R2, "rf_high" = rf_high_R2)
  return(scores)
}
```

```{r}
# Trying shit out

rf_2008 <- tune_random_forest(smear_2008_2011)
gbm_2008 <- tune_gbm(smear_2008_2011)
two_var_2008 <- train_with_two(smear_2008_2011, gbm_2008, rf_2008, 1)

rf_2012 <- tune_random_forest(smear_2012_2019)
gbm_2012 <- tune_gbm(smear_2012_2019)
two_var_2012 <- train_with_two(smear_2012_2019, gbm_2012, rf_2012, 1)
```

```{r, fig.height = 7}
p1 <- plot_for_lowNhigh(two_var_2008[["gbm_low"]], two_var_2008[["gbm_high"]], "GBM 2008-2011")
p2 <- plot_for_lowNhigh(two_var_2008[["rf_low"]],  two_var_2008[["rf_high"]],  "RF 2008-2011")

grid.arrange(p1, p2, nrow=2, heights=c(30, 30))

p1 <- plot_for_lowNhigh(two_var_2012[["gbm_low"]], two_var_2012[["gbm_high"]], "GBM 2012-2019")
p2 <- plot_for_lowNhigh(two_var_2012[["rf_low"]],  two_var_2012[["rf_high"]],  "RF 2012-2019")

grid.arrange(p1, p2, nrow=2, heights=c(30, 30))
```

```{r}
frame <- create_hilo_pred_frame(rf_2008, high_points, low_points)
plot_hi_lo(frame, "RF, Trained With 2008-2011")

frame <- create_hilo_pred_frame(gbm_2008, high_points, low_points)
plot_hi_lo(frame, "GBM, Trained With 2008-2011")

frame <- create_hilo_pred_frame(rf_2012, high_points, low_points)
plot_hi_lo(frame, "RF, Trained With 2008-2019")

frame <- create_hilo_pred_frame(gbm_2012, high_points, low_points)
plot_hi_lo(frame, "GBM, Trained With 2012-2019")
```

```{r}
names(sort(colMeans(two_var_2008[["gbm_low"]]), decreasing = TRUE)[1:3])
names(sort(colMeans(two_var_2008[["gbm_high"]]), decreasing = TRUE)[1:3])

names(sort(colMeans(two_var_2008[["rf_low"]]), decreasing = TRUE)[1:3])
names(sort(colMeans(two_var_2008[["rf_high"]]), decreasing = TRUE)[1:3])

names(sort(colMeans(two_var_2012[["gbm_low"]]), decreasing = TRUE)[1:3])
names(sort(colMeans(two_var_2012[["gbm_high"]]), decreasing = TRUE)[1:3])

names(sort(colMeans(two_var_2012[["rf_low"]]), decreasing = TRUE)[1:3])
names(sort(colMeans(two_var_2012[["rf_high"]]), decreasing = TRUE)[1:3])


```